Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-11-26 Latex_files/Thesis/chapters/chapter 2/main.tex
updates to chp 4 & 2

\chapter{Theory and methods}
\label{chapter:theory}
The two key areas of this project cover two widely different fields; 
crystallisation theory, which is covered in the previous introduction 
section, and tweezing theory. The following chapter summarises the 
working principles behind optical tweezing, the differences in 
scattering theories, and the broader theoretical calculations used
throughout the rest of the Thesis.

@@

Proper understanding of optical tweezing requires an understanding of
how the trapped particle interacts with the trapping laser. From an 
electromagnetism perspective the laser creates a spatially and 
temporally coherent electric field that scatters light off of a 
trapped particle.  The laws governing electric and magnetic fields 
are summarised most succinctly via the Maxwell equations. The 
differential forms of which are given below \cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
current, and $\epsilon_0$ is the permittivity of free space.

@@

The scattering force arises from the scattered field of the trapping 
beam pushing the target object away from the focal point. While some 
trapping arrangements can induce transverse scattering forces for 
simple spheres the scattering force is only significant in the direction
of propagation \cite{Capitanio2002} The equilibrium position is found 
when the gradient force far cancels out the the scattering force, 
and the thermal energy of the particle.

@@

The strength of the optical trap is directly proportional to both 
the laser power and the contrast in refractive indices between the 
medium and the particle. Fully describing and calculating the 
optical forces is dependent on the ratio between the particle size 
($a$) and the trapping wavelength ($\lambda$). The scattering theory 
for computing the optical force is dependent on this ratio.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lornez-Mie Theory}

@@

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves. For a laser of arbitrary shape the 
electric field can be described as a summation of partial 
spherical waves multiplied by an expansion coefficient. Depending 
on the beam in question the expansion coefficients can be tailored
to match the desired beam shape. For a Gaussian beam interacting 
with a particle of volume $V$, the incident, internal, and 
scattered fields are given as \cite{RanhaNeves2019}: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
	E_0 \sum_{n}^\infty\sum_{m=-n}^n \left[
	a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	+ b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
	\\
	\label{eq:internal}
	{\bf E}_{\rm int}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^{n}\left[
	d_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	\right]
	\\
	\label{eq:scattered}
	{\bf E}_{\rm scat}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^n\left[
	p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
	\right] 
\end{align}

@@

In order to compute the optical force and torque applied to a 
given object one simply needs to know the momentum carried by 
the incident and scattered fields. So long as we know the 
expansion coefficients of the incident beam we can find the 
scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to 
compute the force. This project utilises the results from 
Farsund \cite{Farsund1996} but in the form used by Crichton 
\cite{Crichton2000THEMD}. As an example the linear momentum 
carried by any beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

And the linear momentum carried by the scattered beam is given by:
\begin{equation}
	\label{eq:linear_momentum_out}
	\begin{split}
		\langle P_{z,\ out} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[p_{nm}q^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
	\end{split}
\end{equation}

Therefore the total force imparted to the trapped particle is thus:
\begin{align}
	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
	\label{eq:Fz}
\end{align}

\noindent
Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in Sec.~\ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size, 
though as the size of the target particle changes - relative to 
the wavelength of light - the theory describing the electromagnetic 
fields are better easier to approximate by alternative theories. 
When $a \gg \lambda$ the optical force is best described by the 
Ray-Optics model. In the opposite case the target particle is 
better approximated as a single electric dipole as described by 
the Rayleigh approximation.

@@

The Rayleigh approximation is for describing a particle when 
$a \ll \lambda$. The underlying theory is that a dielectric sphere 
can be treated as a dipole while in the presence of the 
electromagnetic field. In which case the scattering force is 
given simply by the scattering of the induced dipole, and the 
gradient force is due to the Lorentz force \cite{Gordon1973}. 
The gradient forces in the principle Cartesian axis are 
described by Harada et al \cite{YasuhiroHarada1996} in MKS units 
as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \label{eq:Rayleigh_start}\\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{z}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \label{eq:Rayleigh_end}
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:Fz}, namely that the 
Rayleigh regime is not concerned with the state of the scattered
field. This is because if the particle is approximated by a point
dipole we can ignore the momentum transfer calculation and simply
compute the Lorenz force for a dipole in an electric field. While 
not important for optical force calculation higher complexity 
scattering problems can be simplified by subdividing the particle 
into discrete dipoles (see Sec~\ref{sec:scattering}), in which case 
the scattered field is now significant. As the target particle gets 
larger this assumption fails to accurately describe the trapping force 
\cite{Li2021}. For larger particles the optical force and scattering 
is better described via the Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered 
field produced by a particle. While one can compute this by 
directly applying Lorenz-Mie theory, describing the interactions 
between focused beams and complicated shapes can be computationally 
slow and memory intensive. Several alternative scattering approaches 
have been developed that allow for faster computation times 
simplifying the scattering problem. While all of these methods will
produce identical results the key difference is in how one represents
the scatterer.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}

The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (ECBM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. By expanding the harmonic functions one can 
derive the individual elements of \textbf{T} until a desired 
accuracy is achieved. One can either use the ECBM to compute the 
scattering for a fixed particle orientation and position. Or they
can evaluate the scattering for any particle orientation by 
computing its $T$-matrix.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker 
\cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. This is essentially an 
extension of the Rayleigh Regime, though in order to compute the 
optical force one requires a full description of the electric 
field. Each dipole interacts with both the incident field and the 
scattered fields from every other dipole surrounding it. The 
resulting scattered field is identical to the scattered field 
produced by direct integration of Eq.~\eqref{eq:internal} 
throughout the full particle volume \cite{Goedecke1988}. The 
integral form for the total electric field inside a scatterer is 
given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  
	\mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

@@

The DDA method can be used to derive the target's $T$-matrix. After
generating the far field expansion of the electric field (Eq.~
\eqref{eq:DDA}), the vector spherical expansion far field 
(Eq.~\eqref{eq:scattered}) can be matched to it. Solving for the 
expansion coefficients $p_{nm}$, and $q_{nm}$ allows one to construct
the T-matrix of the target.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Langevin Equation}
Describing any microscopic motion requires an understanding of a 
molecules diffusive behaviour, for the case of optical tweezers 
the most complete model of diffusion is the Langevin equation. 
Models such as the Fickian, and Einstein derivations are sufficient 
for macroscopic behaviours the Langevin equation better describes 
the microscopic characteristics of any diffusive behaviour.

The Fickian model describes the net flux of solute molecules into a 
finite volume of fluid being proportional to the density gradient 
$\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. It assumes that the 
solute molecules do not collide with one another or with molecules in 
the solution; overall, the Fickian model is used to describe how solute 
molecules disperse over long periods of time. It does not provide any 
insight into the forces acting on individual molecules, whereas the 
Langevin model captures the physical interactions between an individual 
molecule and the surrounding fluid over a wide range of time scales.

The Einstein model of diffusion expands upon the Fickian model by 
considering the collisions between individual molecules. If we 
consider a single particle suspended in a solution it will experience 
multiple collisions with other molecules \cite{Gillespie2012a} over a 
given time interval $\delta t$. The Einstein model allows us to 
consider how individual molecules behave while suspended in a 
surrounding fluid. Where it begins to falter is when we consider 
bringing $\delta t$ to $0$; in this scenario the molecule experiences 
not many but only individual collisions. The Einstein model does not 
consider the inertia of the molecule in question and so for very short 
time frames the model implies that the molecule's velocity changes 
instantly after each collision \cite{Gillespie2012a,Gillespie2012b}. 
Furthermore, the kinetic energy of each collision is not limited by 
the thermal energy of the system, meaning that using the Einstein model 
to predict a particle's trajectory to an finite degree of accuracy 
implies that the particle has infinite kinetic energy \cite{Gillespie2012b}. 
This failure to describe motion over smaller time frames is addressed 
in the Langevin model by accounting for the fluid drag of the system, 
where any sudden change in velocity must result in a proportionally 
opposed drag force \cite{Gillespie2012c}.

@@

Positional data is often acquired in an experimental setting either using
image analysis or photodiodes to infer the particles position relative to
the beam focus. The former is often used in cases when precision is not a 
key concern, as often a standard CCD camera will have limited spacial and 
temporal resolution. The latter method often requires the use of a quadrant
photo diode(QPD), the data provided by the QPD will often need to be converted
to physical units to compute the force directly - see Eq.~\eqref{eq:conversion_factor}.

@@

\subsubsection{Angular Mean Square Displacement (MSAD)}
It is also possible to plot the angular MSD (MSAD) using simulative data, 
however using this to describe the torque acting on an asymmetric object 
is difficult due to fact there is no simple approximation of the rotational
stiffness. Vigilante \textit{et al} \cite{Vigilante2020} derived the upper 
limit of a dimer's MSAD along its long axis by assuming it was strongly 
trapped and so had limited angular motion, there expression gives:
\begin{align}
	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
	2\left[1-\frac{1}{4\beta\kappa_r} 
	\left(\frac{exp(\beta\kappa_r)-1}
	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
	}\right)^2\right]
\end{align}

Where, $u_z$ is the unit vector connecting the two spheres of the dimer, 
$\beta=1/k_BT$, $\kappa_r$ is the rotational stiffness of the trap, and $F$ 
is Dawson's integral \cite{Oldham2008}. Vigilante's paper expressed that 
they couldn't compute $MSD(\tau)$ because they couldn't solve the 
Einstein-Smoluchowski equation which describes the diffusion constant for 
dielectric particles. This would require a full description of a 
particle's electrical mobility and charge distribution - the latter could 
be achieved via a discrete dipole approximation, the former would be 
dependent on both the particle's position and relative orientation to the 
electric field.

@@

\subsubsection{Power spectral analysis for rotating objects}
The scattering from rotating objects can be partially characterised via
the power spectrum density method. Typically papers reporting on the 
rotation rate of a trapped object will collect a power spectrum and look
at the maximum frequency term to determine its rotation rate. However this
often neglects any information on the trapping forces acting on particle
and only looks at the torque applying around one of the particle's primary 
axis. There has only been one notable effort to fully characterise the 
full trapping dynamics on a rotating body, \cite{Yogesha2012} developed 
a theoretical model that relates the power spectrum to both the 
rotational and translational motion simultaneously.

@@

While translational motion have no knock on effects to the QPD signal, 
rotational changes are between the incident and total fields. This results
in rotational motion being biased in the QPD signal - even when collecting
signals from isotropic scatterers. To remedy this an inverse rotation is
applied to the total field.

>>>>

\chapter{Theory and methods}
\label{chapter:theory}
The two key areas of this project cover two widely different
fields; crystallisation theory, which is covered in the 
previous introduction section, and tweezing theory. The 
following chapter summarises the working principles behind 
optical tweezing, the differences in scattering theories, 
and the broader theoretical calculations used throughout the 
rest of the thesis.

@@

Proper understanding of optical tweezing requires a description
of how the trapped particle interacts with the trapping laser. 
From an electromagnetism perspective the laser creates a 
spatially and temporally coherent electric field that scatters 
light off of a trapped particle. The laws governing electric 
and magnetic fields are summarised most succinctly via the 
Maxwell equations. The differential forms of which are given 
below \cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the 
charge current, and $\epsilon_0$ is the permittivity of free 
space.

@@

The scattering force arises from the scattered field of the 
trapping beam pushing the target object away from the focal 
point. While some trapping arrangements can induce transverse 
scattering forces for simple spheres the scattering force is 
only significant in the direction of propagation \cite{Capitanio2002}.
The equilibrium position is found when the gradient force 
cancels out the the scattering force, and the thermal energy 
of the particle.

@@

The strength of the optical trap is directly proportional to 
both the laser power and the contrast in refractive indices 
between the medium and the particle. Fully describing the 
optical forces is dependent on the ratio between the particle 
size ($a$) and the trapping wavelength ($\lambda$). The 
scattering theory that describes the optical force is dependent 
on this ratio.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lornez-Mie Theory}

@@

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves. For a laser of arbitrary shape the 
electric field can be described as a summation of partial 
spherical waves multiplied by an expansion coefficient. Depending 
on the beam in question the expansion coefficients can be tailored
to match the desired beam shape. For a approximate Gaussian beam 
interacting with a particle of volume $V$, the incident, internal, 
and scattered fields are given as \cite{RanhaNeves2019}: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
	E_0 \sum_{n}^\infty\sum_{m=-n}^n \left[
	a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	+ b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
	\\
	\label{eq:internal}
	{\bf E}_{\rm int}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^{n}\left[
	d_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	\right]
	\\
	\label{eq:scattered}
	{\bf E}_{\rm scat}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^n\left[
	p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
	\right] 
\end{align}

@@

In order to compute the optical force and torque applied to a 
given object one simply needs to know the momentum carried by 
the incident and scattered fields. So long as we know the 
expansion coefficients of the incident beam we can find the 
scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to 
compute the force. This project utilises the results from 
Farsund \cite{Farsund1996} but in the form used by Crichton 
\cite{Crichton2000THEMD}. As an example the time averaged linear 
momentum carried by any beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

And the time averaged linear momentum carried by the scattered 
beam is given by:
\begin{equation}
	\label{eq:linear_momentum_out}
	\begin{split}
		\langle P_{z,\ out} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[p_{nm}q^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
	\end{split}
\end{equation}

Therefore the total force imparted to the trapped particle is thus:
\begin{align}
	\boldsymbol{F_z} =\frac{dP}{dt} = \langle P_{z,\ in}\rangle 
	-\langle P_{z,\ out}\rangle
	\label{eq:Fz}
\end{align}

\noindent
Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in Sec.~\ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size, 
though as the size of the target particle changes - relative to 
the wavelength of light - the theory describing the electromagnetic 
fields are easier to approximate by alternative theories. When the 
particle radius is $\gg \lambda$ the optical force is best described 
by the Ray-Optics model. In the opposite case the target particle is 
better approximated as a single electric dipole as described by 
the Rayleigh approximation.

@@

The Rayleigh approximation is for describing a particle when 
the particle radius is $\ll \lambda$. The underlying theory 
is that a dielectric sphere can be treated as a dipole while 
in the presence of the electromagnetic field. In which case 
the scattering force is given simply by the scattering of 
the induced dipole, and the gradient force is due to the 
Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative 
rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \label{eq:Rayleigh_start}\\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{z}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \label{eq:Rayleigh_end}
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:Fz}, namely that the 
Rayleigh regime is not concerned with the state of the scattered
field. This is because if the particle is approximated by a point
dipole we can ignore the momentum transfer calculation and simply
compute the Lorenz force for a dipole in an electric field. While 
not important for optical force calculation higher complexity 
scattering problems can be simplified by subdividing the particle 
into discrete dipoles (see Sec~\ref{sec:scattering}), in which case 
the scattered field is now significant. As the target particle gets 
larger this approximation fails to accurately describe the trapping force 
\cite{Li2021}. For larger particles the optical force and scattering 
is better described via the Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered 
field produced by a particle. While one can compute this by 
directly applying Lorenz-Mie theory, describing the interactions 
between focused beams and complicated shapes can be computationally 
slow and memory intensive. Several alternative scattering approaches 
have been developed that allow for faster computation times 
simplifying the scattering problem. While all of these methods will
produce similar results the key difference is in how one represents
the scatterer.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (ECBM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. By expanding the harmonic functions one can 
derive the individual elements of \textbf{T} until a desired 
accuracy is achieved. One can either use the ECBM to compute the 
scattering for a fixed particle orientation and position. Or they
can evaluate the scattering for any particle orientation by 
computing its $T$-matrix.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker 
\cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. This is essentially an 
extension of the Rayleigh approximation, though in order to compute 
the optical force one requires a full description of the electric 
field. Each dipole interacts with both the incident field and the 
scattered fields from every other dipole surrounding it. The 
resulting scattered field is identical to the scattered field 
produced by direct integration of Eq.~\eqref{eq:internal} 
throughout the full particle volume \cite{Goedecke1988}. The 
integral form for the total electric field inside a scatterer is 
given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  
	\mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

@@

The DDA method can be used to derive the target's $T$-matrix. After
generating the far field expansion of the electric field (Eq.~
\eqref{eq:DDA}), the vector spherical expansion far field 
(Eq.~\eqref{eq:scattered}) can be matched to it. Solving for the 
expansion coefficients $p_{nm}$, and $q_{nm}$ allows one to construct
the T-matrix of the target. 
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Describing the trajectory of a trapped particle}
While describing the optical forces is sufficient for 
theoretical characterisation of an optical trap, trying 
to replicate that model in an experimental setting is 
near impossible. As such it is often far simpler to 
characterise the optical trap by recording the particle's
trajectory (see \ref{sec:position_detection} for 
specifics regarding the equipment), and then fitting it 
to our presumed description of the beam. The trajectory 
is easily understood if we understand the particle's 
diffusive behaviour.

\section{Langevin Equation}
Describing any microscopic motion requires an understanding 
of a molecules diffusive behaviour, for the case of optical 
tweezers the most complete model of diffusion is the Langevin 
equation. Models such as the Fickian, and Einstein derivations 
are sufficient for macroscopic behaviours the Langevin 
equation better describes the microscopic characteristics of 
any diffusive behaviour.

The Fickian model describes the net flux of solute molecules 
into a finite volume of fluid being proportional to the 
density gradient $\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. 
It assumes that the solute molecules do not collide with one 
another or with molecules in the solution; overall, the Fickian 
model is used to describe how solute molecules disperse over 
long periods of time. It does not provide any insight into 
the forces acting on individual molecules, nor does it prove 
particularly useful over smaller time scales. Whereas the Langevin 
model captures the physical interactions between an individual 
molecule and the surrounding fluid over a wide range of time 
scales.

The Einstein model of diffusion expands upon the Fickian model 
by considering the collisions between individual molecules. If 
we consider a single particle suspended in a solution it will 
experience multiple collisions with other molecules 
\cite{Gillespie2012a} over a given time interval $\delta t$. 
Applying statistical analysis to an observed trajectory allows
one to make measurements of a particle's diffusive behaviour 
within a fluid. Where it begins to falter is when we consider 
bringing $\delta t$ to $0$; in this scenario the molecule 
experiences not many but only individual collisions. The 
Einstein model does not consider the inertia of the molecule 
in question and so for very short time frames the model 
implies that the molecule's velocity changes instantly after 
each collision \cite{Gillespie2012a, Gillespie2012b}. 
Furthermore, the kinetic energy of each collision is not 
limited by the thermal energy of the system, meaning that 
using the Einstein model to predict a particle's trajectory 
to an finite degree of accuracy implies that the particle 
has infinite kinetic energy \cite{Gillespie2012b}. This 
failure to describe motion over smaller time frames is 
addressed in the Langevin model by accounting for the fluid 
drag of the system, where any sudden change in velocity must 
result in a proportionally opposed drag force \cite{Gillespie2012c}.

@@

Positional data is often acquired in an experimental setting 
either using image analysis or photodiodes to infer the particles 
position relative tothe beam focus. The former is often used in 
cases when precision is not a key concern, as often a standard 
CCD camera will have limited spacial and temporal resolution. 
The latter method often requires the use of a quadrant photo 
diode(QPD), the data provided by the QPD will often need to be 
converted to physical units to compute the force directly - see Eq.~\eqref{eq:conversion_factor}.

@@

\subsubsection{Angular Mean Square Displacement (MSAD)}
It is also possible to plot the angular MSD (MSAD) using 
simulative or experimental data, but there is no complementary 
expression for \eqref{eq:MSD}. This is because while the 
particle is close to the focus the optical force is easily 
modelled as a Hookean spring; conversely, the optical torque
is far more complex even if the particle is strongly trapped. 
Vigilante \textit{et al} \cite{Vigilante2020} derived the 
upper limit of a dimer's MSAD along its long axis by assuming 
it was strongly trapped and so had limited angular motion, 
there expression gives:
\begin{align}
	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
	2\left[1-\frac{1}{4\beta\kappa_r} 
	\left(\frac{exp(\beta\kappa_r)-1}
	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
	}\right)^2\right]
\end{align}

Where, $u_z$ is the unit vector connecting the two spheres of 
the dimer, $\beta=1/k_BT$, $\kappa_r$ is the rotational stiffness 
of the trap, and $F$ is Dawson's integral \cite{Oldham2008}. 
Vigilante's paper expressed that they couldn't compute 
$MSAD(\tau)$ because they couldn't solve the Einstein-Smoluchowski 
equation which describes the diffusion constant for dielectric 
particles. This would require a full description of a particle's 
electrical mobility and charge distribution - the latter could be 
achieved via a discrete dipole approximation, the former would be 
dependent on both the particle's position and relative orientation 
to the electric field.

@@

\subsubsection{Power spectral analysis for rotating objects}
The scattering from rotating objects can be partially characterised via
the power spectrum density method. If the particle is birefringent then
as it rotates the detected signal will oscillate periodically. When a 
Fourier transform is taken the power spectra will have a peak signal at 
that periodic signal.Typically papers reporting on the rotation rate of 
a trapped object will collect a power spectrum and look at the maximum frequency term to determine its rotation rate. However this often neglects 
any information on the trapping forces acting on particle and only looks 
at the torque applying around one of the particle's primary axis. There 
has only been one notable effort to fully characterise the full trapping dynamics on a rotating body, \cite{Yogesha2012} developed a theoretical 
model that relates the power spectrum to both the rotational and 
translational motion simultaneously. This is applicable to particles with
a polarizable long axis but is insufficient for birefringent particles.

@@

While translational motion has no knock on effects to the QPD signal, 
rotational changes are double counted by both the incident and total 
field. This results in rotational motion being biased in the QPD signal 
- even when collecting signals from isotropic scatterers. To prevent 
this we need to rotate the total field via the inverse rotation matrix
of the dimer. Shown in fig, is a flow diagram outlining the full work
flow of a given trajectory, and how the QPD data is acquired.

\begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{Simulation_workflow.pdf}
	\caption{Block flow diagram of how a trajectory and QPD data is 
		calculated and collected. Relevant equations are mentioned 
		at each step.}
\end{figure}
\newpage
